---
title: "HW1"
author: "Daniel Pontoh"
date: "January 22, 2017"
output: html_document
---
```{r q0, warning=FALSE, message=FALSE}
library(readxl)
library(ggplot2)
library(dplyr)
library(tidyverse)


#standard read
#URL <- "https://data.bls.gov/timeseries/LNU04000000?years_option=all_years&periods_optio#n=specific_periods&periods=Annual+Data"
#fil <- basename(URL)
#fil <- file.path("data", fil)
#
#if (!file.exist(fil)) download.file(URL, fil)
```
####The data was not able to easily download. Go to link above and manually download the .xlsx 

# Question 1
#####Is the current (as close to present as possible with the data) national unemployment rate going up, down or holding steady?
```{r q1, warning=FALSE, message=FALSE}
unemp <- read_excel("data/SeriesReport-20170122150353_c71d1e.xlsx", skip=10, col_names=TRUE)
unemp <- mutate(unemp, Year=as.Date(sprintf("%d-01-01", as.integer(Year))))
```

#####Trend line for all entire time period of the sample:

```{r q1a, warning=FALSE, message=FALSE}
ggplot(unemp, aes(Year, Annual)) +
  geom_line(aes(group=1)) +
  geom_point() +
  stat_smooth() +
  scale_x_date(date_breaks="10 years", date_labels="%Y")
```

#####Trend line for he past 10 years:

```{r q1b, warning=FALSE, message=FALSE}

unemp08on <- filter(unemp, Year > as.character.Date("2007-01-01"))

ggplot(unemp08on, aes(Year, Annual)) +
   geom_line(aes(group=1)) +
   geom_point() +
   stat_smooth() +
   scale_x_date(date_breaks="2 years", date_labels="%Y")
```


# Question 2
######What was the mean unemployment rate for New Jersey by complete decade?
```{r q2, warning=FALSE, message=FALSE}
# New Jersey Mean Unemployment Rate 2006-15 --------

unemp <- read_excel("data/staadata.xlsx")

unemp <- unemp[8:nrow(unemp),c(2,3,10)]
colnames(unemp) <- c("State","Year", "Rate")

unemp$Rate <- as.numeric(unemp$Rate)

unemp05 <- unemp[unemp$Year > 2005,]
njunemp <- unemp05[unemp05$State == "New Jersey",]
a <- mean(njunemp$Rate)
a
```
#####New Jersey's mean unemployment rate for the past decade was 7.2%

# Question 3
######How does the mean unemployment rate for New Jersey by complete decade compare to the mean unemployment rate by complete decade for for Wyoming?
```{r q3}
# NJ vs WY Mean Unemployment Rate 2006-15 --------


wyunemp <- unemp05[unemp05$State == "Wyoming",]
mean(wyunemp$Rate)

mean(njunemp$Rate) - mean(wyunemp$Rate)

```

#####New Jersey's Mean Unemployment Rate for years 2006-15 is 2.61 percentage points greater than Wyoming's for that period of time.

# Question 4
######Which state(s) — could be ties — have the lowest (most current) unemployment rate?

```{r q4}
# State(s) w/ current lowest unemployment rate ----------------------------

unemp15 <- unemp05[unemp05$Year == "2015",]
unemp15[order(unemp15$Rate),] %>%
  head()

unemp15[which(unemp15$Rate == min(unemp15$Rate)),]


```

#####North Dakota had the lowest unemployment rate in 2015.

# Question 5
######Which states(s) — could be ties — had the lowest unemployment unemployment rate in 2008?
```{r q5}
# State(s) w/ lowest unemployment rate in 2008 ----------------------------

unemp08 <- unemp05[unemp05$Year == "2008",]
unemp08[order(unemp08$Rate),] %>%
  head() 

unemp08[which(unemp08$Rate == min(unemp08$Rate)),]

```
#####South Dakota and Wyoming were tied for lowest unemployment rate in 2008.
